---
title: "Portfolio 1"
---

> The projects should be numbered consecutively (i.e., in the order in which you began them), and should include for each project a description of the goal, the product (computer program, hand graph, computer graph, etc.), the data, and some interpretation. Reports must be reproducible and of high quality in terms of writing, grammar, presentation, etc.

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidyr)
```

```{r}
library(tidyverse)
library(dplyr)
library(haven)
Mother_and_teen_data <- read_sav("Mother_and_teen_data.sav")
#View(Mother_and_teen_data)
Mother_and_teen_data$Race3 <- as.factor(Mother_and_teen_data$Race3)
select(Mother_and_teen_data, PSE, CD_1, CD_2, CD_3, PAPWARM, PAPmoni, PAPsolic, PAPrule, Race3)
ggplot(Mother_and_teen_data, aes(x=PSE, y=CD_1, fill = Race3)) +  geom_smooth(size=1, alpha=0.4) +
 geom_point(aes(color=Race3))+
  labs(title="PSE and adolescents' disobeying parents", x="Parental self-efficacy", y="Disobeying parents among adolescents" ) 
  
```

```{r}
ggplot(Mother_and_teen_data, aes(x=PSE, y=CD_2, fill = Race3)) +  geom_smooth(size=1, alpha=0.4) +
 geom_point(aes(color=Race3))+
  labs(title="PSE and adolescents' school misconduct", x="Parental self-efficacy", y="school misconduct among adolescents" ) 
  
```

```{r}
ggplot(Mother_and_teen_data, aes(x=PSE, y=CD_3, fill = Race3)) +  geom_smooth(size=1, alpha=0.4) +
 geom_point(aes(color=Race3))+
  labs(title="PSE and adolescents' substance use", x="Parental self-efficacy", y="substance use among adolescents" ) 
```
Trying to combine three type of adolescent deviance
```{r}
Mother_and_teen_data$Race3 <- as.factor(Mother_and_teen_data$Race3) 
Mother_and_teen_data <- Mother_and_teen_data %>%
mutate(ad_deviance=CD_1 + CD_2 + CD_3)
ggplot(Mother_and_teen_data, aes(x=PSE, y=ad_deviance)) +  geom_smooth(size=1, alpha=0.4) +
 geom_point(aes(color=Race3)) +
  labs(title="PSE and adolescents' substance use", x="Parental self-efficacy", y="adolescent deviance" ) 
```
SEM to test the model and its fits
```{r} 
model1<-"AdolD=~CD_1 +CD_2 +CD_3
EffectP=~PAPmoni+PAPWARM+PAPsolic+PAPrule
PSE~EffectP
AdolD~EffectP + PSE"
#residual correlations
PAPmoni~~PAPsoli
PAPmoni~~PAPrule
PAPWARM~~PAPsoli
CD_1~~CD_2
CD_1~~CD_3
CD_2~~CD_3
```
the model fits.
```{r} 
library(lavaan)
library(readxl)
library(dplyr)
library(haven)
teendata<- Mother_and_teen_data %>% filter( !is.na(PerPW)) 
fit1<- sem(model1, data=Mother_and_teen_data)
fit1
summary(fit1, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit1, sort.=TRUE)
inspect(fit1, what="std")
```