---
title: "Portfolio 10"
---

> The projects should be numbered consecutively (i.e., in the order in which you began them), and should include for each project a description of the goal, the product (computer program, hand graph, computer graph, etc.), the data, and some interpretation. Reports must be reproducible and of high quality in terms of writing, grammar, presentation, etc.

The ten portfolios concentrated on building up graphs and data analyses for my first year and thesis projects. I used SEM, ggplot, multigroup modeling, t-test, ANOVA tests, and chi-square tests to analyze my data and build up graphs for my presentations on the conference. Basically I tried to redo the data analyses that I had done in SPSS for my first year project, and then did some data analyses for my thesis. R did a great job at visualizing the associations and interactions that I aimed to investigate in my thesis.


```{r make graphs}
#for the poster presentation, make a graph that show the effect
library(tidyverse)
library(dplyr)
library(haven)
Mother_and_teen_data <- read_sav("Mother_and_teen_data.sav")
#View(Mother_and_teen_data)
Mother_and_teen_data$Race3 <- as.factor(Mother_and_teen_data$Race3)
```
```{r build models}
library(tidyverse)
library(dplyr)
library(haven)
mt <- Mother_and_teen_data %>% 
 filter(!is.na(Race3)) %>% 
  mutate(ad_deviance=CD_1 + CD_2 + CD_3)
model1<-"AdolD=~CD_1 +CD_2 +CD_3
EffectP=~PAPmoni+PAPWARM+PAPsolic+PAPrule
PSE~EffectP
AdolD~EffectP + PSE"
#residual correlations
PAPmoni~~PAPsoli
PAPmoni~~PAPrule
PAPWARM~~PAPsoli
CD_1~~CD_2
CD_1~~CD_3
CD_2~~CD_3
```
the model fits.
```{r model..} 
library(lavaan)
library(readxl)
library(dplyr)
library(haven)
teendata<- Mother_and_teen_data %>% filter( !is.na(PerPW)) 
fit1<- sem(model1, data=Mother_and_teen_data)
fit1
summary(fit1, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit1, sort.=TRUE)
inspect(fit1, what="std")

```
#Then I tested multiple models with different intervariable correlations.
#try to let all the effective parenting factors be able to correlate with each other
```{r test model}
model2<-"AdolD=~CD_1 +CD_2 +CD_3
EffectP=~PAPmoni+PAPWARM+PAPsolic+PAPrule
PSE~EffectP
AdolD~EffectP + PSE"
#residual correlations
PAPmoni~~PAPsoli
PAPmoni~~PAPrule
PAPWARM~~PAPsoli
PAPWARM~~PAPmoni
PAPWARM~~PAPrule
PAPsoli~~PAPrule

CD_1~~CD_2
CD_1~~CD_3
CD_2~~CD_3

library(lavaan)
library(readxl)
library(dplyr)
library(haven)
teendata<- Mother_and_teen_data %>% filter( !is.na(PerPW)) 
fit2<- sem(model2, data=Mother_and_teen_data)
fit2
summary(fit2, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit2, sort.=TRUE)
inspect(fit2, what="std")
fitmeasures(fit2)

#does not make any noticeable change.
```
```{r test model 3 }
#try to let let monitoring correlates with the rules and solicitation correlates with monitoring
model3<-"AdolD=~CD_1 +CD_2 +CD_3
EffectP=~PAPmoni+PAPWARM+PAPsolic+PAPrule
PSE~EffectP
AdolD~EffectP + PSE"
#residual correlations
PAPmoni~~PAPsoli
PAPmoni~~PAPrule


CD_1~~CD_3

library(semPlot)
library(lavaan)
library(readxl)
library(dplyr)
library(haven)
teendata<- Mother_and_teen_data %>% filter( !is.na(PerPW)) 
fit3<- sem(model3, data=Mother_and_teen_data)
fit3
summary(fit3, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit3, sort.=TRUE)
inspect(fit3, what="std")

parameterestimates(fit3, standardized = TRUE)
fitted(fit3)
residuals(fit3)
fitmeasures(fit3)
modificationindices(fit3, sort. = TRUE)
semPaths(fit3, what="paths", whatLables="par", rotation=2)

fitmeasures(fit3)

#does not make any noticeable change as well.
#did not find significant better model. all three model seems alike. The fit measure difference is unnoticeable.
```
```{r comparison}

library(ggpubr)

compare_means(PSE ~ Race3,  data = mt)
my_comparisons <- list( c("1", "2"), c("1", "3"), c("2", "3") )
ggboxplot(mt, x = "Race3", y = "PSE",
          color = "Race3", palette = "jco")+ 
  labs(title= "Cultural Difference in PSE", x="Race/Ethnicity", y="PSE", color="Race/Ethnicity") + 
  scale_x_discrete(labels=c('White', 'Latina', 'African American'))+
  stat_compare_means(comparisons = my_comparisons)+ 
  stat_compare_means(label.y = 7)    

#VERY SATISFIED WITH THIS PLOT!
```

```{r model greenish}
test_model1<-"AdolD=~CD_1 +CD_2 +CD_3
EffectP=~PAPmoni+PAPWARM+PAPsolic+PAPrule
PSE~EffectP
AdolD~EffectP + PSE"
#residual correlations
PAPmoni~~PAPsoli
PAPmoni~~PAPrule
PAPWARM~~PAPsoli

CD_1~~CD_2
CD_1~~CD_3
CD_2~~CD_3
library(semPlot)
library(lavaan)
library(readxl)
library(dplyr)
library(haven)
teendata<- Mother_and_teen_data %>% filter( !is.na(PerPW)) 
fit1<- sem(model1, data=Mother_and_teen_data)
fit1
summary(fit1, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit1, sort.=TRUE)
inspect(fit1, what="std")

parameterestimates(fit1, standardized = TRUE)
fitted(fit1)
residuals(fit1)
fitmeasures(fit1)
modificationindices(fit1, sort. = TRUE)
semPaths(fit1, what="paths", whatLables="par", rotation=2)

fitmeasures(fit1)

#looks like the model made by r is uglier than what I got from Amos. I might be stuck with Amos output.
```
```{r result}
semPaths(fit1,"std",layout = 'tree', edge.label.cex=.9, curvePivot = TRUE)
library(tidyverse)
library(knitr)
library(lavaan)
library(psych)
library(MBESS)
```
```{r}
library(tidyverse)
library(dplyr)
library(haven)
Mother_and_teen_data <- read_sav("Mother_and_teen_data.sav")
#View(Mother_and_teen_data)
mt <- Mother_and_teen_data %>% 
 filter(!is.na(Race3)) %>% 
  mutate(ad_deviance=CD_1 + CD_2 + CD_3)
mt$Race3 <- as.factor(mt$Race3)
moderation <- lm(ad_deviance ~ PSE * Race3, mt)
summary(moderation)
moddp <- lm(CD_1 ~ Race3 * PSE, mt)
summary(moddp)

ggplot(mt, aes(PSE, CD_1, color = factor(Race3))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Race/Ethnicity", labels = c("European Americans", "Latinx", "African Americans"), values = c("#FF6666", "#6666FF", "green")) + 
  labs(x="PSE", y="Disobeying Parents") +
  theme_classic()

```
```{r disobeying parents}
# try to test whether race/ethnicity moderates the association between PSE and Disobeying parents (CD_1)
moddp <- lm(CD_1 ~ Race3 * PSE, mt)
summary(moddp)

ggplot(mt, aes(PSE, CD_1, color = factor(Race3))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Race/Ethnicity", labels = c("European Americans", "Latinx", "African Americans"), values = c("#FF6666", "#6666FF", "green")) + 
  labs(x="PSE", y="Disobeying Parents") +
  theme_classic()

#non-significant as well.
```
```{r school misconduct}
# try to test whether race/ethnicity moderates the association between PSE and school misconduct (CD_2)
modsm <- lm(CD_2 ~ Race3 * PSE, mt)
summary(modsm)

ggplot(mt, aes(PSE, CD_2, color = factor(Race3))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Race/Ethnicity", labels = c("European Americans", "Latinx", "African Americans"), values = c("#FF6666", "#6666FF", "green")) + 
  labs(x="PSE", y="School Misconduct") +
  theme_classic()

#non-significant as well.
```
```{r substance use}
# try to test whether race/ethnicity moderates the association between PSE and substance uses (CD_3)
modsu <- lm(CD_3 ~ PSE * Race3, mt)
summary(modsu)

ggplot(mt, aes(PSE, CD_3, color = factor(Race3))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Race/Ethnicity", labels = c("European Americans", "Latinx", "African Americans"), values = c("#FF6666", "#6666FF", "green")) + 
  labs(x="PSE", y="Substance Use") +
  theme_classic()

#Race/ethnicity significantly moderates the association between PSE and substance use. PSE negatively associates with substance use for European Americans, but not for the Latinx.
#same results as what I got from SPSS.
```
```{r data}
library(tidyverse)
library(dplyr)
library(haven)
teen_survey <- read_sav("teen_survey.sav")
```
#use my thesis data. 
```{r create composites}
teen_survey$exter<- rowMeans(teen_survey[,c("Q8_1", "Q8_2", "Q8_3", "Q8_4", "Q8_5", "Q8_6", "Q8_7", "Q8_8", "Q8_9")],
                             na.rm = TRUE)
teen_survey$inter<- rowMeans(teen_survey[,c("Q9_1", "Q9_2", "Q9_3", "Q9_4", "Q9_5", "Q9_6", "Q9_7", "Q9_8", "Q9_9", "Q9_10", "Q9_11", "Q9_12", "Q9_13", "Q9_14")],
                             na.rm = TRUE)
teen_survey$stem<- rowMeans(teen_survey[,c("Q10_1", "Q10_2", "Q10_3", "Q10_4", "Q10_5", "Q10_6", "Q10_7", "Q10_8", "Q10_9", "Q10_10", "Q10_11")],
                             na.rm = TRUE)
teen_survey$pros<- rowMeans(teen_survey[,c("Q12_1", "Q12_2", "Q12_3", "Q12_4", "Q12_5", "Q12_6", "Q12_7", "Q12_8")],
                             na.rm = TRUE)
teen_survey$girl<- rowMeans(teen_survey[,c("Q14_1", "Q14_2", "Q14_3", "Q14_4", "Q14_5")],
                             na.rm = TRUE)
teen_survey$boy<- rowMeans(teen_survey[,c("Q18_1", "Q18_2", "Q18_3", "Q18_4", "Q18_5")],
                             na.rm = TRUE)
teen_survey$ex<- rowMeans(teen_survey[,c("Q20_1", "Q20_2", "Q20_3", "Q20_4", "Q20_5", "Q20_6", "Q20_7", "Q20_8", "Q20_9")],
                             na.rm = TRUE)
teen_survey$int<- rowMeans(teen_survey[,c("Q21_1", "Q21_2", "Q21_3", "Q21_4", "Q21_5", "Q21_6", "Q21_7", "Q21_8", "Q21_9", "Q21_10", "Q21_11", "Q21_12", "Q21_13", "Q21_14")],
                            na.rm = TRUE)
teen_survey$st<- rowMeans(teen_survey[,c("Q22_1", "Q22_2", "Q22_3", "Q22_4", "Q22_5", "Q22_6", "Q22_7", "Q22_8", "Q22_9", "Q22_10", "Q22_11")],
                             na.rm = TRUE)
teen_survey$pr<- rowMeans(teen_survey[,c("Q24_1", "Q24_2", "Q24_3", "Q24_4", "Q24_5", "Q24_6", "Q24_7", "Q24_8")],
                             na.rm = TRUE)
teen_survey$appe<- rowMeans(teen_survey[,c("Q25_1", "Q25_2", "Q25_3", "Q25_4", "Q25_5", "Q25_6", "Q25_7", "Q25_8", "Q25_9", "Q25_10")],
                             na.rm = TRUE)
teen_survey$ind<- rowMeans(teen_survey[,c("Q26_1", "Q26_2", "Q26_3", "Q26_4", "Q26_5", "Q26_6")],
                             na.rm = TRUE)
teen_survey$dep<- rowMeans(teen_survey[,c("Q27_1", "Q27_2", "Q27_3", "Q27_4", "Q27_5", "Q27_6")],
                             na.rm = TRUE)
teen_survey$fam<- rowMeans(teen_survey[,c("Q30_1", "Q30_2", "Q30_3", "Q30_4", "Q30_5", "Q30_6")],
                             na.rm = TRUE)
teen_survey$piety<- rowMeans(teen_survey[,c("Q31_1", "Q31_2", "Q31_3", "Q31_4", "Q31_5", "Q31_6", "Q31_7")],
                             na.rm = TRUE)
teen_survey$gen<- rowMeans(teen_survey[,c("Q32_1", "Q32_2", "Q32_3", "Q32_4", "Q32_5", "Q32_6", "Q32_7", "Q32_8", "Q32_9")],
                             na.rm = TRUE)

```

```{r factor}
teen_survey$country <- as.factor(teen_survey$country)
teen_survey$Gender <- as.factor(teen_survey$Gender)
```
```{r t_package}

library(lessR)
```
```{r mutate}
ts <- teen_survey %>%
  filter(Gender %in% c("1", "2"))
```
#try to use multilevel but it was not correct. the coding was correct for the function though.
```{r multilevel ex}
library(tidyverse)
library(dplyr)
library(haven)
teen_survey<- read_sav("teen_survey.sav")
tss<- read_sav("tss.sav")
library(lme4)
library(merTools)
library(semPlot)
library(lavaan)
library(readxl)
library(dplyr)
library(haven)
model1<-lmer(ex~1+(1|exter),REML = FALSE, data=tss)
summary(model1)
confint(model1)

ICC(outcome = "ex", group="exter",data=tss)
#add level 1 predictor
model2<-lmer(ex~1+Gender+(1|exter), REML= FALSE, data= tss)
summary(model2)
anova(model1, model2)
confint(model2)
model3<-lmer(ex~1+Gender+(1|exter)+country, REML= FALSE, data= tss)
summary(model3)
anova(model2, model3)
confint(model3)
#does not think it make any sense statistically.
```
```{r build sem model for thesis}

model2<-"Parents_expectations=~exter +inter+ pros+ stem
Adolescents_behaviors=~int+ex+pr+st
Parents_expectations~Adolescents_behaviors"
#residual correlations
exter~~inter
pros~~stem

int~~ex
pr~~int
pr~~ex
st~~pr
st~~ex
st~~int




library(lavaan)
library(readxl)
library(dplyr)
library(haven)

fit2<- sem(model2, data=teen_survey)

summary(fit2, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit2, sort.=TRUE)
inspect(fit2, what="std")
fitmeasures(fit2)


parameterestimates(fit2, standardized = TRUE)
fitted(fit2)
residuals(fit2)
fitmeasures(fit2)
modificationindices(fit2, sort. = TRUE)
semPaths(fit2, what="paths", whatLables="par", rotation=2)

fitmeasures(fit2)

#poor fit.
```
```{r third try of sem}
teen_survey_1<- read_sav("teen_survey_1.sav")
gender<-read_sav("gender.sav")
#rerun model with the updated dataset
model4<-"Parents_expectations=~exter +inter+ pros+ stem
Adolescents_behaviors=~int+ex+pr+st
Parents_expectations~Adolescents_behaviors"
#residual correlations
exter~~inter
pros~~stem

int~~ex
pr~~int
pr~~ex
st~~pr
st~~ex
st~~int




library(lavaan)
library(readxl)
library(dplyr)
library(haven)

fit4<- sem(model4, data=teen_survey_1)

summary(fit4, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit4, sort.=TRUE)
inspect(fit4, what="std")
fitmeasures(fit4)


parameterestimates(fit4, standardized = TRUE)
fitted(fit4)
residuals(fit4)
fitmeasures(fit4)
modificationindices(fit4, sort. = TRUE)
semPaths(fit4, what="paths", whatLables="par", rotation=2)

fitmeasures(fit4)

#still not a good fit
```

```{r data new}
teen_survey_1<- read_sav("teen_survey_1.sav")
```
```{r graphs}
ggplot(teen_survey_1, aes(x=exter, y=ex, color = factor(country))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Country", labels = c("Americans", "Chineses"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived Expectations for Externalizing", y="Externalizing") +
  theme_classic()
ggplot(teen_survey_1, aes(x=inter, y=int, color = factor(country))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Country", labels = c("Americans", "Chineses"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived Expectations for Internalizing", y="Internalizing") +
  theme_classic()
#looks like the perceived parents' expectations for internalizing predict actual internalizing for Chinese more compared to U.S. adolescents.
ggplot(teen_survey_1, aes(x=stem, y=st, color = factor(country))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Country", labels = c("Americans", "Chineses"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived Expectations for STEM", y="STEM") +
  theme_classic()
#looks like the perceived parents' expectations for STEM motivations predict actual STEM motivations for Chinese more compared to U.S. adolescents.
#also, the Chinese reported lower perceived expectations for STEM motivations and adolescents' actual STEM motivations.

```

```{r prosocial}
ggplot(teen_survey_1, aes(x=pros, y=pr, color = factor(country))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Country", labels = c("Americans", "Chineses"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived Expectations for prosocial behaviors", y="Prosocial behaviors") +
  theme_classic()
```
```{r ex}
#visualize gender difference
ggplot(gender, aes(x=exter, y=ex, color = factor(Gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Gender", labels = c("Girls", "Boys"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived Expectations for externalizing", y="externalizing") +
  theme_classic()
```
```{r internalizing}
ggplot(gender, aes(x=inter, y=int, color = factor(Gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Gender", labels = c("Girls", "Boys"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived Expectations for internalizing", y="internalizing") +
  theme_classic()
```
```{r prosocial..}
ggplot(gender, aes(x=pros, y=pr, color = factor(Gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Gender", labels = c("Girls", "Boys"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived Expectations for prosocial behaviors", y="prosocial behaviors") +
  theme_classic()
```
```{r stem.}
ggplot(gender, aes(x=stem, y=st, color = factor(Gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Gender", labels = c("Females", "Males"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived Expectations for STEM", y="STEM") +
  theme_classic()

```
```{r viausalize appe}
ggplot(gender, aes(x=APPE, y=st, color = factor(Gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Gender", labels = c("Females", "Males"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived parental efficacy", y="STEM motivations") +
  theme_classic()
ggplot(gender, aes(x=APPE, y=ex, color = factor(Gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Gender", labels = c("Females", "Males"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived parental efficacy", y="externalizing") +
  theme_classic()
ggplot(gender, aes(x=APPE, y=int, color = factor(Gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Gender", labels = c("Females", "Males"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived parental efficacy", y="internalizing") +
  theme_classic()
ggplot(gender, aes(x=APPE, y=pr, color = factor(Gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Gender", labels = c("Females", "Males"), values = c("#FF6666", "#6666FF")) + 
  labs(x="Perceived parental efficacy", y="prosocial behaviors") +
  theme_classic()

#seems like there were interactions between APPE and gender in intenalizing, externalizing, and prosocial behaviors. 
#girls internalizing and prosocial behavirs were all negatively impacted by APPE, which is interesting. I wonder why that is the case.

```
#miss the fact that gender and country are dependent of each other and it was hard to tell that the effects were from gender or country difference without controlling.
```{r}
table(gender$Gender,gender$country)
chisq.test(gender$Gender, gender$country, correct=FALSE)
```
